MSDS6372Project1

### 2021-01-27 MSDS6372Project1.Rmd
adjusting models for objective 1 with VIF and Summary for selecting the best predictors for simple model.

```{r}

#### Objective 2 start without Log transformation - since there are so many categorical variables

>>>>

```{r}
# Explore numerical variables
train %>% select(MSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity)%>%ggpairs()
# Try Log transformation on MSRP see if it will improve correlations
train$logMSRP = log(train$MSRP)
train %>% select(logMSRP,Year,EngineHP,EngineCylinders,NumberOfDoors,HighwayMPG,CityMPG,Popularity)%>%ggpairs()
# It looks like logMSRP has increased correlation with Year but decreased correlation with EngineHP, EngineCylinders
# It's best to run LASSO and choose predictors on both with transformation and without Log transformation in order to identify which
# Model performs better
train = train[,-27]
#str(train)
library(glmnet)
library(car)
# Since we want a easy to interperate model, we don't want to use predictor Make and Model that has too many levels in the model
train = train[,-(1:2)]
str(train)
#Dummy code categorical predictor variables
x <- model.matrix(MSRP~.,train)[,-1]
y <- log(train$MSRP)
set.seed(123) # Need a seed number to keep same results for LASSO
cv.out=cv.glmnet(x,y,alpha=1) #alpha=1 performs LASSO
par(mfrow=c(1,1))
plot(cv.out)
# Look at coefficients
coef(cv.out, cv.out$lambda.min)
# Seems like LASSO has selected all predictors, only few levels within EngineFuel, TransmissionType, VehicleStyle got turned off
# Build model with predictors LASSO selected
log.model1 = lm(log(MSRP)~.,data=train)
# Check VIF
summary(log.model1)
# TransmissionType and Hatchback provides NA on summary statistic, try remove them as predictors
log.model2 = lm(log(MSRP)~.- TransmissionType - Hatchback,data=train)
summary(log.model2)
# Look at VIF
vif(log.model2)[,3]^2
# CityMPG has really high VIF remove it from model
log.model3 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG,data=train)
# Look at VIF
vif(log.model3)[,3]^2
# NumberOfDoors has really high VIF remove from model
log.model4 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors,data=train)
# Look at VIF
vif(log.model4)[,3]^2
# HighwayMPG still have over 10 VIF, remove from model
log.model5 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG,data=train)
vif(log.model5)[,3]^2
# Diesel is not significant, HighPerformance is not significant, Performance is not significant Remove from simple model
log.model6 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG - Diesel - HighPerformance - Performance,data=train)
vif(log.model6)[,3]^2
# N_A is still not significant, remove from model
log.model7 = lm(log(MSRP)~.- TransmissionType - Hatchback - CityMPG - NumberOfDoors - HighwayMPG - Diesel - HighPerformance - Performance - N_A,data=train)
vif(log.model7)[,3]^2
summary(log.model7)
# Model 7 is a highly interpretable model with decent R-squared.

par(mfrow=c(2,2))
plot(log.model7)
library(olsrr)
ols_plot_cooksd_bar(log.model1)
ols_plot_resid_stand(log.model1)
# Residual looks okay~
# Now try MSRP without log transformation

#Dummy code categorical predictor variables
x1 <- model.matrix(MSRP~.,train)[,-1]
y1 <- train$MSRP
set.seed(123) # Need a seed number to keep same results for LASSO
cv.out1=cv.glmnet(x1,y1,alpha=1) #alpha=1 performs LASSO
par(mfrow=c(1,1))
plot(cv.out1)
# Look at coefficients
coef(cv.out1, cv.out1$lambda.min)
# Again, all predictors are selected, only few levels are turned off
regular.model1 = lm(MSRP~.,data=train)
summary(regular.model1)
# Delete Hatchback & TransmissionType... NA on summary table
regular.model2 = lm(MSRP~.-Hatchback-TransmissionType,data=train)
vif(regular.model2)[,3]^2
# remove CityMPG - highest VIF, larger than 29
regular.model3 = lm(MSRP~.-Hatchback-TransmissionType-CityMPG,data=train)
vif(regular.model3)[,3]^2
# remove NumberOfDoors
regular.model4 = lm(MSRP~.-Hatchback-TransmissionType-CityMPG-NumberOfDoors,data=train)
vif(regular.model4)[,3]^2
# HighwayMPG still over 10
regular.model5 = lm(MSRP~.-Hatchback-TransmissionType-CityMPG-NumberOfDoors-HighwayMPG,data=train)
vif(regular.model5)[,3]^2
summary(regular.model5)
# FlexFuel has the highest P value in terms of significance - delete
regular.model6 = lm(MSRP~.-Hatchback-TransmissionType-CityMPG-NumberOfDoors-HighwayMPG-FlexFuel,data=train)
vif(regular.model6)[,3]^2
summary(regular.model6)
#Diesel is high in p-value delete
regular.model7 = lm(MSRP~.-Hatchback-TransmissionType-CityMPG-NumberOfDoors-HighwayMPG-FlexFuel-Diesel,data=train)
vif(regular.model7)[,3]^2
summary(regular.model7)
par(mfrow=c(2,2))
plot(regular.model1) # Residual plot doesn't look as well as log(MSRP) residual so tranformation is necessary.
ols_plot_cooksd_bar(regular.model1)
ols_plot_resid_stand(regular.model1)

#############################################################################
# By comparison Constant variance looks a lot better with the log(MSRP)...###
# Since the assumption works best with this, next check with stepwise method to see what predictors will be selected
stepwise.model1 =lm(log(MSRP)~.,data=train)
ols_step_both_aic(stepwise.model1)
# By Using StepwiseAIC variable selection method, the following predictors are selected
# Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+TransmissionType+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel
log.model2.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+TransmissionType+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel,data=train)
summary(log.model2.step)
# TransmissionType has NA need to be deleted
log.model3.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+CityMPG+Popularity+N_A+Diesel,data=train)
summary(log.model3.step)
vif(log.model3.step)[,3]^2
# CityMPG has the highest VIF - delete
log.model4.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+HighwayMPG+Hybrid+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A+Diesel,data=train)
vif(log.model4.step)[,3]^2
# HighwayMPG still over 10, delete
log.model5.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+FlexFuel+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A+Diesel,data=train)
vif(log.model5.step)[,3]^2
summary(log.model5.step)
#FlexFuel has the highest insignificant P-value delete
log.model6.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A+Diesel,data=train)
summary(log.model6.step)
#Diesel has the highest insignificant P-value delete
log.model7.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity+N_A,data=train)
summary(log.model7.step)
#N_A has the highest insignificant P-value delete
log.model8.step = lm(log(MSRP)~Year+EngineHP+Exotic+EngineFuelType+VehicleStyle+Hybrid+EngineCylinders+Luxury+VehicleSize+DrivenWheels+NumberOfDoors+Crossover+FactoryTuner+Popularity,data=train)
summary(log.model8.step)

par(mfrow=c(2,2))
plot(log.model8.step)
ols_plot_cooksd_bar(log.model8.step)
ols_plot_resid_stand(log.model8.step)

# LASSO and STEPWISE came to slight different Model in terms of predictor and r-squared value.
# LASSO r-squared 0.8426, adjusted r-squared 0.8419
# Stepwise r-squared 0.843, adjusted r-squared 0.8424
#compare test ASE results between stepwise model and LASSO
test = test[,-(1:2)]
testMSRP = test$MSRP # GET real result 
test = test[,-13] # Delete MSRP from test 
# Calculate stepwise test ASE first
pred.test.step = predict(log.model8.step,test)
# exp values to return to original value
pred.test.step = exp(pred.test.step)
test_ASE_STEP = mean((testMSRP-pred.test.step)^2)
test_ASE_STEP #560738819

#Calculate LASSO test ASE
pred.test.lasso = predict(log.model7,test)
pred.test.lasso = exp(pred.test.lasso)
test_ASE_LASSO = mean((testMSRP-pred.test.lasso)^2)
test_ASE_LASSO #571497373

# By comparison, Stepwise model workflow provided the best model in terms of r-squared, adjusted r-squared and test ASE.
# Move forward with the analysis with log.model8.step
```

# Note: Rerun the code chunk for clean train/test/validate datasets - altered in Objective 1
#### Objective 2 start without Log transformation - since there are so many categorical variables
